---
title: "Calculate recovery indicators"
author: "Wanda De Keersmaecker"
date: ""
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Calculate recovery indicators}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  ---
  
```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = TRUE, eval=F)
```

# Load required libraries
```{r}
library(raster)
library(ff)
library(UpscaleRecovery)
```

# Inputs
```{r}
ifolder <- 'C:\\Users\\keers001\\OneDrive\ -\ WageningenUR\\RETURN\\Data\\RETURN\\20200107_Upscaling\\CaseStudy\\Data\\' # directory with NBR data
ofolder <- 'C:\\Users\\keers001\\OneDrive\ -\ WageningenUR\\RETURN\\Data\\RETURN\\20200107_Upscaling\\CaseStudy\\Data\\' # directory where the recovery indicators and temporary outputs will be saved
firefolder <- 'C:\\Users\\keers001\\OneDrive - WageningenUR\\RETURN\\Data\\RETURN\\CCI_fire\\' #'\\\\WURNET.NL\\Homes\\keers001\\AppData\\FolderRedirection\\Desktop\\cci_fire2\\' 
# directory where the fire data are stored

tsfile <- 'LSTS_SantaRem_area2_2000_2020'# name of the file with NBR time series
metafile <- 'LSTS_meta_SantaRem_area2_2000_2020'# metadata file of the NBR time series
firefile <- 'LSTS_FireArea_SantaRem_area2_2000_2020'# name of the file with the fire mask

tempRes <- 'quart' # daily, annual, monthly, or quart (yearly data not supported yet!)
tempFun <- 'max' # max or mean (function used for temporal aggregation of data)
timeThres <- 2 # in case of segmentations, the time period between the detected break and the disturbance should be less than timeThres years
slpThres <- 2 # the slope of the pre-disturbance period should be less than slpThres

# settings of the derived recovery indicators
# input: refers to the tine series pre-processing: raw (=no preprocessing), smoothed or segmented
# h: This parameter defines the minimal segment size either given as fraction relative to the sample size or as an integer giving the minimal number of observations in each segment.
# shortDenseTS: TRUE or FALSE. In case TRUE, the metrics are adjusted to be compatible with short, dense time series. In case FALSE, the input time series is assumed to have annual observations and at least 2 and 5 pre- and post-disturbance years, respectively.
# nPre: If shortDenseTS is TRUE, number of years prior to the disturbance used to calculate the pre-disturbance value
# nDist: If shortDenseTS is TRUE, number of months used to quantify the time series value during the disturbance
# nPostMin: If shortDenseTS is TRUE, min number of years after the disturbance used to quantify the recovery
# nPostMax: If shortDenseTS is TRUE, max number of years after the disturbance used to quantify the recovery
# tdist: the timing of the disturbance [observation number]

if((tempRes == 'monthly') || (tempRes == 'daily')){
  funSet <- list('input' = c('raw', 'smoothed', 'segmented', 'raw', 'smoothed', 'segmented'),# settings for the recovery indicators
                 'shortDenseTS' = c( TRUE, TRUE, TRUE, TRUE, TRUE, TRUE),
                 'nPre' = c(2,2,2,2,2,2),
                 'nDist' = c(1,1,1,12,12,12),
                 'nPostMin' = c(0,0,0,4,4,4),
                 'nPostMax' = c(1,1,1,6,6,6),
                 'h' = c(0.15,0.15,0.15,0.15,0.15,0.15))
} else if (tempRes == 'annual'){
  funSet <- list('input' = c('raw'),# settings for the recovery indicators
                 'shortDenseTS' = c( FALSE),
                 'nPre' = c(2),
                 'nDist' = c(12),
                 'nPostMin' = c(4),
                 'nPostMax' = c(6),
                 'h' = c(0.15))
}

```

# Import data
```{r}
br <- stack(paste0(ifolder,tsfile,'.tif'))# the data stack containing the vegetation response
br <- br[[-1]] # the first image is redundant, artefact from the GEE download script
meta <- read.csv(paste0(ifolder,metafile,'.csv'))# metadata associated with the image stack
dts <- as.Date(meta$system_time_start[-1],'%Y,%m,%d')# observation dates of the image stack
fre <- raster(paste0(ifolder,firefile,'.tif'))# raster of areas that experienced a fire (those have a value > 0)
fcl <- loadRData( file = file.path(firefolder, 'ESACCI-L3S_FIRE-BA-MODIS-AREA_2-fv5.1-CL-01-18.rda'))# fire confidence
fjd <- loadRData( file = file.path(firefolder, 'ESACCI-L3S_FIRE-BA-MODIS-AREA_2-fv5.1-JD-01-18.rda'))# fire day of year (doy)
```

# Resampling fire data at ca 30m spatial resolution
General preprocessing of the data: change the spatial resolution and extent
```{r}
# crop and change spatial resolution of fire stack
fcl30 <- crop(fcl,extent(br))# change extent of the image stack
fcl30 <- resample(fcl30, br, method="ngb")#resample

fjd30 <- crop(fjd,extent(br))
fjd30 <- resample(fjd30, br, method="ngb")

# # Locations that experienced fire 
freloc <- (fre > 0) # areas that have experienced a fire get the value 1, other areas the value 0

# date vector associated with the fire stack
fdts<- as.Date(names(fcl), format = "X%Y%m%d") # dates associated with the fire data stack

# generate an image stack containing regular fire time series at the predefined temporal resolution with value 1 if a fire occured and 0 if no fire occured
tsFire <- calc(stack(freloc,fcl30, fjd30), function(x){createFireStack(x, dts = fdts, resol = tempRes, thres = 95)}, filename = paste0(ofolder, tsfile, '_Fire_', tempRes),overwrite=TRUE)

```

# Preprocess NBR data
Create regular time series
```{r}
# get the dates of the time series observations
names(br) <- dts
dts <- as.Date(dts, format = "X%Y.%m.%d") ## needed as input in the helper function of get_m_agg

# Create regular time series 
tsbr <- calc(stack(freloc,br), function(x){toRegularTSStack(x, dts, fun = tempFun, resol = tempRes)}, filename = paste0(ofolder, tsfile, '_', tempRes, '_', tempFun), overwrite=TRUE)


```

# Preprocess fire data 
The fire data are converted to the right temporal resolution and its representation is adjusted (value 0 represents no fire and 1 a fire)
```{r}
# e <-  drawExtent()
# fst <- crop(stack(freloc,fcl30,fjd30),e)

# convert the fire data to a regular time series with same temporal resolution as the NBR series. Here, observations during fire events get the value 1 and other observations the value 0

tsFire <- calc(stack(freloc,fcl30,fjd30), function(x){createFireStack(x, fdts, resol = tempRes, thres = 95)}, filename = paste0(ofolder, tsfile, '_Fire_', tempRes), overwrite=TRUE)

# get the dates of the nbr and fire stacks
if(tempRes == 'monthly'){
  dtsbr <- seq(min(dts), max(dts), by = "1 month")#"1 day"
  dtsbr <- rollback(dtsbr, roll_to_first = TRUE, preserve_hms = TRUE)
  dtsfr <- fdts
  dtsfr <- rollback(dtsfr, roll_to_first = TRUE, preserve_hms = TRUE)
}
save(dtsfr, file = file.path(ofolder, paste0(tsfile, '_Firedts_', tempRes)))
save(dtsbr, file = file.path(ofolder, paste0(tsfile, '_dts_', tempRes, '_', tempFun)))

```

# Plot time series
Here, an example of a time series with fire is plotted
```{r}
# load time series
tsbr <- stack(paste0(ofolder, tsfile, '_', tempRes, '_', tempFun, '.gri'))
tsFire <- stack(paste0(ofolder, tsfile, '_Fire_', tempRes, '.gri'))

# find a pixel that experienced a fire
fireind <- Which(freloc>0, cells = T)
ii <- fireind[20]

# get the time series from the stack 
tsii <- as.numeric(tsbr[ii])
frii <- as.numeric(tsFire[ii])

# plot time series
plot(zoo(tsii,dtsbr))
lines(zoo(frii,dtsfr), col = 'red')

```

# Create total stack

Make sure the the fire and NBR stack cover the same time period and merge both stacks
```{r}
# ensure that the two stacks cover the same time period 

# get the start and end date of the overlap period
startdt <- max(c(min(dtsbr), min(dtsfr)))
enddt  <- min(c(max(dtsbr), max(dtsfr)))

# clip the stacks to the overlap period
indbr  <- which((dtsbr > (startdt-1)) & (dtsbr < (enddt+1)))
indfr  <- which((dtsfr > (startdt-1)) & (dtsfr < (enddt+1)))

finbr <- tsbr[[indbr]]
finfr <- tsFire[[indfr]]

# # combine mask, fire and nbr stacks into one stack
st <- stack(freloc, finfr, finbr)
```

# Calculate recovery indicators
Four recovery indicators are computed: the RRI, R80P, YrYr and the slope of the post-disturbance segment
```{r}
# calculate stability indicators
# RRI,R80P, YrYr, Sl

for(i in 1:length(funSet[[1]])){# iterate over the settings to derive the recovery indicators
  # get the correct number of observations per year
   if(tempRes == 'monthly') {
    obspyr = 12} else if(tempRes == 'daily'){obspyr = 365}else if(tempRes == 'annual'){obspyr = 1}
  
  # calculate the recovery indicators
  out <- calc(st, function(x){calcRecoveryStack(x, maxBreak=F, obspyr=obspyr, inp = funSet$input[i], shortDenseTS = funSet$shortDenseTS[i], nPre = funSet$nPre[i], nDist = funSet$nDist[i], nPostMin = funSet$nPostMin[i],
                                                nPostMax = funSet$nPostMax[i], h = funSet$h[i], timeThres, slpThres)}, 
              filename = paste0(ofolder, tsfile, '_Stab_', tempRes, '_maxBreak_F', '_obspyr_', obspyr, '_inp_', funSet$input[i], '_shortDenseTS_', funSet$shortDenseTS[i], '_nPre_',  funSet$nPre[i], '_nDist_', funSet$nDist[i], '_nPostMin_', funSet$nPostMin[i], '_nPostMax_' , funSet$nPostMax[i], '_h_', funSet$h[i]*100, '_', timeThres, '_', slpThres), overwrite=TRUE)
}

```

```{r, include=FALSE}
stab <- stack(file.path(ofolder,'LSTS_SantaRem_area2_2000_2020_Stab_monthly_maxBreak_F_obspyr_12_inp_segmented_shortDenseTS_TRUE_nPre_2_nDist_1_nPostMin_0_nPostMax_1_h_15.gri'))
plot(stab[[1]])
```

