---
title: "Plot recovery indicators"
author: "Wanda De Keersmaecker"
date: ""
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Plot recovery indicators}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval= F)
```
# Load libraries
```{r}
library(UpscaleRecovery)
library(ggplot2)
library(raster)
library(rasterVis)
library(sp)
library(maps)
library(maptools)
library(ggmap)
library("ggplot2")
library("sf")
library('rgdal')
library(rnaturalearth)
library(rnaturalearthdata)
library("ggspatial")
library("RColorBrewer")
```

```{r}
# folder where data is stored
ofolder <- 'C:\\Users\\keers001\\OneDrive - WageningenUR\\RETURN\\Data\\RETURN\\20200107_Upscaling\\CaseStudy\\Data\\'
# base name of the recovery indicator files
basename <- c('LSTS_SantaRem_area', '_2000_2020_Stab_monthly_maxBreak_F_obspyr_12_inp_segmented_shortDenseTS_TRUE_nPre_2_nDist_12_nPostMin_4_nPostMax_6_h_15')
# directory where the figures will be stored
figfolder <- 'C:\\Users\\keers001\\OneDrive - WageningenUR\\RETURN\\Data\\RETURN\\20200107_Upscaling\\CaseStudy\\Figures\\'

# lcfolder <- 'C:\\Users\\keers001\\OneDrive - WageningenUR\\RETURN\\Data\\RETURN\\CopernicusLandCover\\'
```

# Load the recovery indicators for each site
```{r}
map.world <- map_data('world')

# load the recovery indicator raster stacks
stab1 <- stack(file.path(ofolder, paste0(basename[1],1, basename[2],'.gri')))
stab2 <- stack(file.path(ofolder, paste0(basename[1],2, basename[2],'.gri')))
stab3 <- stack(file.path(ofolder, paste0(basename[1],3, basename[2],'.gri')))
stab4 <- stack(file.path(ofolder, paste0(basename[1],4, basename[2],'.gri')))
```

# Generate point coordinates of the site centroids
```{r}
# generate point coordinates of site centroids
(sites <- data.frame(longitude = c( mean(extent(stab1)[1:2]),  mean(extent(stab2)[1:2]), mean(extent(stab3)[1:2]),  mean(extent(stab4)[1:2])), latitude = c(mean(extent(stab1)[3:4]), mean(extent(stab2)[3:4]),mean(extent(stab3)[3:4]),  mean(extent(stab4)[3:4]))))
(sites <- st_as_sf(sites, coords = c("longitude", "latitude"), 
    crs = 4326, agr = "constant"))

```

# overview of sampled sites
```{r}
theme_set(theme_bw())

world <- ne_countries(scale = "medium", returnclass = "sf")
# fr <- raster('C:\\Users\\keers001\\OneDrive - WageningenUR\\RETURN\\Data\\RETURN\\CopernicusLandCover\\W060N00_ProbaV_LC100_epoch2015_global_v2.0.1_forest-type-layer_EPSG-4326.tif')
# frp <- rasterToPoints(fr); frp <- data.frame(hdf)
# colnames(frp) <- c("X","Y","Forest_type")


plt <- ggplot(data = world) +
  theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue")
  )+
    geom_sf() +
  # layer(geom="raster",data=frp,mapping=aes(X,Y,fill=Forest_type))+
  # scale_fill_gradientn(name="Forest type",colours = rainbow(20))+
  # guides(fill = guide_colorbar()) +
  # geom_polygon(data=forest, aes(x=long, y=lat, group=group), fill="forestgreen", 
  #       colour="forestgreen", alpha=0.8)+
    geom_sf(data = sites, size = 4, shape = 23, fill = "darkred") +
    # annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0, "in"), pad_y = unit(0, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = c(-30, -85), ylim = c(-30, 10), expand = FALSE)
  # png(file.path(figfolder, 'sites.png'))
  # print(plt)
  # dev.off()
  
  ggsave(
  file.path(figfolder, 'sites.png'),
  plt,
  width = 3.25,
  height = 3.25,
  dpi = 1200
)

```

# plot recovery indicators of each site

```{r}
recnames <- c('RRI', 'R80P', 'YrYr', 'Slope')

for (ste in 2:4){
stabste <- switch(ste, stab1, stab2, stab3, stab4)
lctile <- switch(ste, 'W060N00', 'W060N00', 'W060N00', 'W060N00')
xlm <- switch(ste, c(-55.35, -55.29), c(-54.86242, -54.7716), c(-56.34263, -56.24694), c(-52.7285, -52.63175))
ylm <- switch (ste,  c(-2.97,-2.895), c(-3.05422, -2.980127), c(-2.621285, -2.521258), c(-6.163072, -6.053118))

# generate fire background layer
f1 <- raster(file.path(ofolder, paste0('LSTS_FireArea_SantaRem_area',ste,'_2000_2020.tif')))
f1 <- (f1 > 0)
fp1 <- rasterToPolygons(f1,fun=function(x){x==1}, dissolve=TRUE)
fpo1 <- st_as_sf(fp1)

# generate forest background layer
forest <- raster(file.path(lcfolder, paste0(lctile,  '_ProbaV_LC100_epoch2015_global_v2.0.1_forest-type-layer_EPSG-4326.tif')))
forest <- crop(forest, extent(f1))
forestp1 <- rasterToPolygons(forest,fun=function(x){x>-1}, dissolve=TRUE)
forestpo1 <- st_as_sf(forestp1)

# iterate over the recovery indicators
for(si in 1:4){
  # prepare recovery indicator data for plotting - convert from raster to dataframe
  s1p <- rasterToPoints(stabste[[si]]); s1p <- data.frame(s1p)
  colnames(s1p) <- c("X","Y","Stab")
  
  # plot the data
  pl <- ggplot(data = world) +
    theme(
      panel.background = element_rect(fill = "lightblue",
                                      colour = "lightblue"))+
    geom_sf() +
    geom_sf(data = forestpo1, colour = "forestgreen", fill = "forestgreen") +
    geom_sf(data = fpo1, colour = "orange", fill = "orange") +
    geom_tile(data = s1p, aes(X,Y,fill = Stab))  +
    scale_fill_gradientn(name=recnames[si],colours = brewer.pal(n = 11, name = 'PiYG'))+
    guides(fill = guide_colorbar()) +
    # annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "br", which_north = "true",
                           pad_x = unit(0, "in"), pad_y = unit(0, "in"),
                           style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = xlm, ylim = ylm, expand = FALSE)
  
  png(file.path(figfolder, paste0(basename[1],ste,basename[2],'_',recnames[si],'.png')))
  print(pl)
  dev.off()
}
}
```

# plot time series associated with a gradient in recovery capacity per site
```{r}
library(strucchange)
library(bfast)
library(lubridate)
firefolder <- 'C:\\Users\\keers001\\OneDrive - WageningenUR\\RETURN\\Data\\RETURN\\CCI_fire\\'

for (ste in 2:4){# sites 
  st <- stack(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_monthly_max.gri')))
  fr <- stack(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_Fire_monthly.gri')))
  stab <- switch(ste,stab1,stab2,stab3,stab4)
  
  metafile <- paste0('LSTS_meta_SantaRem_area',ste,'_2000_2020')

meta <- read.csv(paste0(ofolder,metafile,'.csv'))# metadata associated with the image stack
dts <- as.Date(meta$system_time_start[-1],'%Y,%m,%d')# observation dates of the image stack
load( file = file.path(firefolder, 'ESACCI-L3S_FIRE-BA-MODIS-AREA_2-fv5.1-CL-01-18.rda'))# fire
fdts<- as.Date(names(clst), format = "X%Y%m%d") # dates associated with the fire data stack

dtsbr <- seq(min(dts), max(dts), by = "1 month")#"1 day"
  dtsbr <- rollback(dtsbr, roll_to_first = TRUE, preserve_hms = TRUE)
  dtsfr <- fdts
  dtsfr <- rollback(dtsfr, roll_to_first = TRUE, preserve_hms = TRUE)
  
  # get the start and end date of the overlap period
  startdt <- max(c(min(dtsbr), min(dtsfr)))
  enddt  <- min(c(max(dtsbr), max(dtsfr)))
  dtsst <- seq(startdt, enddt, by = "1 month")#"1 day"
  
  # clip the stacks to the overlap period
  indbr  <- which((dtsbr > (startdt-1)) & (dtsbr < (enddt+1)))
  indfr  <- which((dtsfr > (startdt-1)) & (dtsfr < (enddt+1)))
  
  st <- st[[indbr]]
  fr <- fr[[indfr]]
  
  
  for (si in 1:4){# recovery indicators
    qu <- as.numeric(quantile(stab[[si]] , probs = c(0.001,0.25,0.5,0.75,0.999), na.rm = T))
    tsii <- matrix(NA,5,dim(st)[3])
    frii <- matrix(NA,5,dim(fr)[3])
    trii <- matrix(NA,5,dim(st)[3])
    for (ii in 1:5){# levels of the recovery index
      minind <- Which(abs(stab[[si]] - qu[ii]) == minValue(abs(stab[[si]] - qu[ii])), cells=T)
      tsii[ii,] <- as.numeric(st[minind[1]])
      frii[ii,] <- as.numeric(fr[minind[1]])
      
      tmp <- ts(tsii[ii,], frequency = 12)
      datapp <- bfastpp(tmp, order = 1, lag = NULL, slag = NULL,
                        na.action = na.omit, stl = 'none')
      bp <- breakpoints(response ~ trend, data = datapp, h = 0.15)
      # Extract BFAST trend component and breaks
      cf <- coef(bp)
      tbp <- bp$breakpoints #observation number of break
      indna <- which(is.na(tmp)==F)
      tbp <- indna[tbp]   # correct observation number for missing values
      #Derive trend component without missing values
      bpf <- c(0, tbp, length(tmp))
      trf <- rep(NA,length(tmp))
      for(ti in 1:(length(bpf)-1)){
        trf[(bpf[ti]+1):bpf[ti+1]] <- cf[ti,1] + ((cf[ti,2]*((bpf[ti]+1):bpf[ti+1])))
      }
      trii[ii,] <- trf
    }
    
    zoots <- zoo(t(tsii),dtsst)
    zoofr <- zoo(t(frii),dtsst)
    
    save(zoots, file = file.path(ofolder,paste0('TS_site_',ste, '_',recnames[si])))
    save(zoofr, file = file.path(ofolder,paste0('TSFire_site_',ste, '_',recnames[si])))
    
    
    png(file.path(figfolder, paste0(basename[1],ste, basename[2],'_T',recnames[si],'.png')),
  width     = 5.25,
  height    = 7.25,
  units     = "in",
  res       = 1200,
  pointsize = 15)
    par(mfrow=c(5,1),mar=c(1,6,1,1))
    for (ii in 1:5){
      plot(zoo(tsii[ii,],dtsst), ylab = 'NBR')
      abline(v=dtsst[frii[ii,]==1], lwd = 2, col = 'red')
      # lines(zoo(frii[ii,],dtsst), col = 'red')
      lines(zoo(trii[ii,],dtsst), col = 'darkgreen')
      # grid()
      # text(diff(par("usr")[1:2])/10, sum(par("usr")[3:4])/2, labels="Test") 
      text(grconvertX(0.15,"npc"), grconvertY(0.2, "npc"), paste0(recnames[si], " = ", round(qu[ii], digits = 4)))
      
    }
    dev.off()
    
  }
}
# plot(d$y,type='l',lwd=3,col='red')
# > plot(d$x1,type='l',lwd=1.5)
# > plot(d$x2,type='l',lwd=1.5)
# > plot(d$x3,type='l',lwd=1.5)

# calcRecoveryTS(c(frii[5,],tsii[5,]), maxBreak=F, obspyr=12, inp = 'segmented', shortDenseTS = TRUE,
#                          nPre = 2, nDist = 12, nPostMin = 4, nPostMax = 6, h = 0.15)
# stab3[minind]

```
```{r}
# 
ifolder <- 'C:\\Users\\keers001\\OneDrive\ -\ WageningenUR\\RETURN\\Data\\RETURN\\20200107_Upscaling\\CaseStudy\\Data\\' # folder with data characteristics
firefolder <- 'C:\\Users\\keers001\\OneDrive - WageningenUR\\RETURN\\Data\\RETURN\\CCI_fire\\' #'\\\\WURNET.NL\\Homes\\keers001\\AppData\\FolderRedirection\\Desktop\\cci_fire2\\'

metafile <- 'LSTS_meta_SantaRem_area3_2000_2020'

meta <- read.csv(paste0(ifolder,metafile,'.csv'))# metadata associated with the image stack
dts <- as.Date(meta$system_time_start[-1],'%Y,%m,%d')# observation dates of the image stack
load( file = file.path(firefolder, 'ESACCI-L3S_FIRE-BA-MODIS-AREA_2-fv5.1-CL-01-18.rda'))# fire
fdts<- as.Date(names(clst), format = "X%Y%m%d") # dates associated with the fire data stack

library(lubridate)
dtsbr <- seq(min(dts), max(dts), by = "1 month")#"1 day"
  dtsbr <- rollback(dtsbr, roll_to_first = TRUE, preserve_hms = TRUE)
  dtsfr <- fdts
  dtsfr <- rollback(dtsfr, roll_to_first = TRUE, preserve_hms = TRUE)


```

```{r}
ste <- 2
stm <- stack(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_monthly_max.gri')))
frm <- stack(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_Fire_monthly.gri')))
dtsm <- loadRData(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_dts_monthly_max')))
dtsfm <- loadRData(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_Firedts_monthly')))
stq <- stack(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_quart_max.gri')))
frq <- stack(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_Fire_quart.gri')))
dtsq <- loadRData(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_dts_quart_max')))
dtsfq <- loadRData(file.path(ofolder, paste0('LSTS_SantaRem_area',ste,'_2000_2020_Firedts_quart')))

# get the start and end date of the overlap period
  startdt <- max(c(min(dtsm), min(dtsfm)))
  enddt  <- min(c(max(dtsm), max(dtsfm)))
  dtsst <- seq(startdt, enddt, by = "1 month")#"1 day"
  
  # clip the stacks to the overlap period
  indbr  <- which((dtsm > (startdt-1)) & (dtsm < (enddt+1)))
  indfr  <- which((dtsfm > (startdt-1)) & (dtsfm < (enddt+1)))
  
  stm <- stm[[indbr]]
  frm <- frm[[indfr]]
  
  # get the start and end date of the overlap period
  startdt <- max(c(min(dtsq), min(dtsfq)))
  enddt  <- min(c(max(dtsq), max(dtsfq)))
  dtsst <- seq(startdt, enddt, by = "1 month")#"1 day"
  
  # clip the stacks to the overlap period
  indbr  <- which((dtsq > (startdt-1)) & (dtsq < (enddt+1)))
  indfr  <- which((dtsfq > (startdt-1)) & (dtsfq < (enddt+1)))
  
  stq <- stq[[indbr]]
  frq <- frq[[indfr]]
  
  
  

stab <- stack(file.path(ofolder, paste0(basename[1],ste, basename[2],'.gri')))

si <- 1

qu <- as.numeric(quantile(stab[[si]] , probs = c(0.001,0.25,0.5,0.75,0.999), na.rm = T))
    tsmii <- matrix(NA,5,dim(stm)[3])
    frmii <- matrix(NA,5,dim(frm)[3])
    trmii <- matrix(NA,5,dim(stm)[3])
    tsqii <- matrix(NA,5,dim(stq)[3])
    frqii <- matrix(NA,5,dim(frq)[3])
    trqii <- matrix(NA,5,dim(stq)[3])
    
    for (ii in 1:5){# levels of the recovery index
      minind <- Which(abs(stab[[si]] - qu[ii]) == minValue(abs(stab[[si]] - qu[ii])), cells=T)
      tsmii[ii,] <- as.numeric(stm[minind[1]])
      frmii[ii,] <- as.numeric(frm[minind[1]])
      
      tmp <- ts(tsmii[ii,], frequency = 12)
      datapp <- bfastpp(tmp, order = 1, lag = NULL, slag = NULL,
                        na.action = na.omit, stl = 'none')
      bp <- breakpoints(response ~ trend, data = datapp, h = 0.15)
      # Extract BFAST trend component and breaks
      cf <- coef(bp)
      tbp <- bp$breakpoints #observation number of break
      indna <- which(is.na(tmp)==F)
      tbp <- indna[tbp]   # correct observation number for missing values
      #Derive trend component without missing values
      bpf <- c(0, tbp, length(tmp))
      trf <- rep(NA,length(tmp))
      for(ti in 1:(length(bpf)-1)){
        trf[(bpf[ti]+1):bpf[ti+1]] <- cf[ti,1] + ((cf[ti,2]*((bpf[ti]+1):bpf[ti+1])))
      }
      trmii[ii,] <- trf
      
      tsqii[ii,] <- as.numeric(stq[minind[1]])
      # frqii[ii,] <- as.numeric(frq[minind[1]])
      
      tmp <- ts(tsqii[ii,], frequency = 4)
      datapp <- bfastpp(tmp, order = 1, lag = NULL, slag = NULL,
                        na.action = na.omit, stl = 'none')
      bp <- breakpoints(response ~ trend, data = datapp, h = 0.15)
      # Extract BFAST trend component and breaks
      cf <- coef(bp)
      tbp <- bp$breakpoints #observation number of break
      indna <- which(is.na(tmp)==F)
      tbp <- indna[tbp]   # correct observation number for missing values
      #Derive trend component without missing values
      bpf <- c(0, tbp, length(tmp))
      trf <- rep(NA,length(tmp))
      for(ti in 1:(length(bpf)-1)){
        trf[(bpf[ti]+1):bpf[ti+1]] <- cf[ti,1] + ((cf[ti,2]*((bpf[ti]+1):bpf[ti+1])))
      }
      trqii[ii,] <- trf
    }
    

    plot(zoo(tsmii[1,],dtsm), type = 'o')
    lines(zoo(trmii[1,],dtsm), col = 'red')
    lines(zoo(tsqii[1,],dtsq), col = 'green', type = 'o')
    lines(zoo(trqii[1,],dtsq), col = 'blue')
  
```
