---
title: "Calculate recovery indicators"
author: "Wanda De Keersmaecker"
date: ""
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Calculate recovery indicators}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  ---
  
```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = TRUE, eval=F)
```
# General
This vignette assumes that a data cube with ARD (level2) Landsat (and/or S2) data has been built using FORCE. In addition a mask to select disturbed pixels (have value 1) and fire data (data have value 1 if a fire is present and a value 0 if no fire occurred) are present.

## Load required libraries
```{r}
library(raster)
# library(ff)
library(UpscaleRecovery)
library(stars)
```









<!-- # Inputs -->
<!-- ```{r} -->
<!-- ifolder <- 'C:\\Users\\keers001\\OneDrive\ -\ WageningenUR\\RETURN\\Data\\RETURN\\20200107_Upscaling\\CaseStudy\\Data\\' # directory with NBR data -->

<!-- tempRes <- 'monthly' # daily, annual, monthly, or quart (yearly data not supported yet!) -->
<!-- tempFun <- 'max' # max or mean (function used for temporal aggregation of data) -->
<!-- timeThres <- 2 # in case of segmentations, the time period between the detected break and the disturbance should be less than timeThres years -->
<!-- slpThres <- 2 # the slope of the pre-disturbance period should be less than slpThres -->

<!-- if((tempRes == 'monthly') || (tempRes == 'daily')){ -->
<!--   funSet <- list('input' = c('raw', 'smoothed', 'segmented', 'raw', 'smoothed', 'segmented'),# settings for the recovery indicators -->
<!--                  'shortDenseTS' = c( TRUE, TRUE, TRUE, TRUE, TRUE, TRUE), -->
<!--                  'nPre' = c(2,2,2,2,2,2), -->
<!--                  'nDist' = c(1,1,1,12,12,12), -->
<!--                  'nPostMin' = c(0,0,0,4,4,4), -->
<!--                  'nPostMax' = c(1,1,1,6,6,6), -->
<!--                  'h' = c(0.15,0.15,0.15,0.15,0.15,0.15)) -->
<!-- } else if (tempRes == 'annual'){ -->
<!--   funSet <- list('input' = c('raw'),# settings for the recovery indicators -->
<!--                  'shortDenseTS' = c( FALSE), -->
<!--                  'nPre' = c(2), -->
<!--                  'nDist' = c(12), -->
<!--                  'nPostMin' = c(4), -->
<!--                  'nPostMax' = c(6), -->
<!--                  'h' = c(0.15)) -->
<!-- } -->

<!-- ``` -->


<!-- # Preprocess NBR data -->
<!-- Create regular time series -->
<!-- ```{r} -->
<!-- # get the dates of the time series observations -->
<!-- names(br) <- dts -->
<!-- dts <- as.Date(dts, format = "X%Y.%m.%d") ## needed as input in the helper function of get_m_agg -->

<!-- # Create regular time series  -->
<!-- tsbr <- calc(stack(freloc,br), function(x){toRegularTSStack(x, dts, fun = tempFun, resol = tempRes)}, filename = paste0(ofolder, tsfile, '_', tempRes, '_', tempFun), overwrite=TRUE) -->


<!-- ``` -->

<!-- # Preprocess fire data  -->
<!-- The fire data are converted to the right temporal resolution and its representation is adjusted (value 0 represents no fire and 1 a fire) -->
<!-- ```{r} -->
<!-- # e <-  drawExtent() -->
<!-- # fst <- crop(stack(freloc,fcl30,fjd30),e) -->

<!-- # convert the fire data to a regular time series with same temporal resolution as the NBR series. Here, observations during fire events get the value 1 and other observations the value 0 -->

<!-- tsFire <- calc(stack(freloc,fcl30,fjd30), function(x){createFireStack(x, fdts, resol = tempRes, thres = 95)}, filename = paste0(ofolder, tsfile, '_Fire_', tempRes), overwrite=TRUE) -->

<!-- # get the dates of the nbr and fire stacks -->
<!--   dtsbr <- as.Date(toRegularTS(dts, dts, fun='max', resol = tempRes)) -->
<!--   dtsbr <- rollback(dtsbr, roll_to_first = TRUE, preserve_hms = TRUE) -->
<!--   dtsfr <- as.Date(toRegularTS(fdts, fdts, fun='max', resol = tempRes)) -->
<!--   dtsfr <- rollback(dtsfr, roll_to_first = TRUE, preserve_hms = TRUE) -->

<!-- save(dtsfr, file = file.path(ofolder, paste0(tsfile, '_Firedts_', tempRes))) -->
<!-- save(dtsbr, file = file.path(ofolder, paste0(tsfile, '_dts_', tempRes, '_', tempFun))) -->

<!-- ``` -->

<!-- # Plot time series -->
<!-- Here, an example of a time series with fire is plotted -->
<!-- ```{r} -->
<!-- # load time series -->
<!-- tsbr <- stack(paste0(ofolder, tsfile, '_', tempRes, '_', tempFun, '.gri')) -->
<!-- tsFire <- stack(paste0(ofolder, tsfile, '_Fire_', tempRes, '.gri')) -->

<!-- # find a pixel that experienced a fire -->
<!-- fireind <- Which(freloc>0, cells = T) -->
<!-- ii <- fireind[20] -->

<!-- # get the time series from the stack  -->
<!-- tsii <- as.numeric(tsbr[ii]) -->
<!-- frii <- as.numeric(tsFire[ii]) -->

<!-- # plot time series -->
<!-- plot(zoo(tsii,dtsbr)) -->
<!-- lines(zoo(frii,dtsfr), col = 'red') -->

<!-- ``` -->

<!-- # Create total stack -->

<!-- Make sure the the fire and NBR stack cover the same time period and merge both stacks -->
<!-- ```{r} -->
<!-- # ensure that the two stacks cover the same time period  -->

<!-- # get the start and end date of the overlap period -->
<!-- startdt <- max(c(min(dtsbr), min(dtsfr))) -->
<!-- enddt  <- min(c(max(dtsbr), max(dtsfr))) -->

<!-- # clip the stacks to the overlap period -->
<!-- indbr  <- which((dtsbr > (startdt-1)) & (dtsbr < (enddt+1))) -->
<!-- indfr  <- which((dtsfr > (startdt-1)) & (dtsfr < (enddt+1))) -->

<!-- finbr <- tsbr[[indbr]] -->
<!-- finfr <- tsFire[[indfr]] -->

<!-- # # combine mask, fire and nbr stacks into one stack -->
<!-- st <- stack(freloc, finfr, finbr) -->
<!-- ``` -->

<!-- # Calculate recovery indicators -->
<!-- Four recovery indicators are computed: the RRI, R80P, YrYr and the slope of the post-disturbance segment -->
<!-- ```{r} -->
<!-- # calculate stability indicators -->
<!-- # RRI,R80P, YrYr, Sl -->

<!-- for(i in 1:length(funSet[[1]])){# iterate over the settings to derive the recovery indicators -->
<!--   # get the correct number of observations per year -->
<!--    if(tempRes == 'monthly') { -->
<!--     obspyr = 12} else if(tempRes == 'daily'){obspyr = 365}else if(tempRes == 'annual'){obspyr = 1} -->

<!--   # calculate the recovery indicators -->
<!--   out <- calc(st, function(x){calcRecoveryStack(x, maxBreak=F, obspyr=obspyr, inp = funSet$input[i], shortDenseTS = funSet$shortDenseTS[i], nPre = funSet$nPre[i], nDist = funSet$nDist[i], nPostMin = funSet$nPostMin[i], -->
<!--                                                 nPostMax = funSet$nPostMax[i], h = funSet$h[i], timeThres, slpThres)},  -->
<!--               filename = paste0(ofolder, tsfile, '_Stab_', tempRes, '_maxBreak_F', '_obspyr_', obspyr, '_inp_', funSet$input[i], '_shortDenseTS_', funSet$shortDenseTS[i], '_nPre_',  funSet$nPre[i], '_nDist_', funSet$nDist[i], '_nPostMin_', funSet$nPostMin[i], '_nPostMax_' , funSet$nPostMax[i], '_h_', funSet$h[i]*100, '_', timeThres, '_', slpThres), overwrite=TRUE) -->
<!-- } -->

<!-- ``` -->

<!-- ```{r, include=FALSE} -->
<!-- stab <- stack(file.path(ofolder,'LSTS_SantaRem_area2_2000_2020_Stab_monthly_maxBreak_F_obspyr_12_inp_segmented_shortDenseTS_TRUE_nPre_2_nDist_1_nPostMin_0_nPostMax_1_h_15.gri')) -->
<!-- plot(stab[[1]]) -->
<!-- ``` -->

